---
title: "Predictive model for FIFA 2019"
output: 
  html_document:
    toc: true
---

# Read data

Read the data from the kaggle website https://www.kaggle.com/karangadiya/fifa19.

```{r loadData}
(load("fifa19small.rda"))
rownames(fifa19small) <- fifa19small$Name

fifa19small["R. Lewandowski",]
```


# Feature engineering

Value is skewed. Will be much easier to model sqrt(Value).

```{r featureEng}
fifa19small$SqrtValue <- sqrt(fifa19small$Value.EUR)

fifa19small <- fifa19small[,-c(1, 2, 3, 4, 6)]
```

# Exploration

Value is skewed. Will be much easier to model sqrt(Value).

```{r dataExploration}
library("ggplot2")
library("DALEX")

ggplot(fifa19small, aes(Age, SqrtValue)) +
  geom_point() + geom_smooth(se = FALSE) +
  theme_drwhy()

ggplot(fifa19small, aes(Age)) +
  geom_histogram() +
  theme_drwhy()

ggplot(fifa19small, aes(Reactions, SqrtValue)) +
  geom_point() + geom_smooth(se = FALSE) +
  theme_drwhy()

ggplot(fifa19small, aes(Reactions)) +
  geom_histogram() +
  theme_drwhy()

ggplot(fifa19small, aes(BallControl, SqrtValue)) +
  geom_point() + geom_smooth(se = FALSE) +
  theme_drwhy()

ggplot(fifa19small, aes(BallControl)) +
  geom_histogram() +
  theme_drwhy()

ggplot(fifa19small, aes(ShortPassing, SqrtValue)) +
  geom_point() + geom_smooth(se = FALSE) +
  theme_drwhy()

ggplot(fifa19small, aes(ShortPassing)) +
  geom_histogram() +
  theme_drwhy()

ggplot(fifa19small, aes(Dribbling, SqrtValue)) +
  geom_point() + geom_smooth(se = FALSE) +
  theme_drwhy()

ggplot(fifa19small, aes(Dribbling)) +
  geom_histogram() +
  theme_drwhy()

```

# Create models

Let's create following models:

- `gbm` model with 250 trees 4 levels depth,
- `gbm` model with 250 trees 1 level depth,
- `randomForest` model with 250 trees.

```{r createModels, war}
library("gbm")
fifa_gbm_deep <- gbm(SqrtValue~., data = fifa19small, n.trees = 250, interaction.depth = 4)

fifa_gbm_shallow <- gbm(SqrtValue~., data = fifa19small, n.trees = 250, interaction.depth = 1)

library("ranger")
fifa_rf <- ranger(SqrtValue~., data = fifa19small, num.trees = 250)

library("rms")
fifa_ols <- ols(SqrtValue ~ rcs(Age) + rcs(Special) + rcs(International.Reputation) + rcs(Skill.Moves) + rcs(Crossing) + rcs(Finishing) + rcs(HeadingAccuracy) + rcs(ShortPassing) + rcs(Volleys) + rcs(Dribbling) + rcs(Curve) + rcs(FKAccuracy) + rcs(LongPassing) + rcs(BallControl) + rcs(Acceleration) + rcs(SprintSpeed) + rcs(Agility) + rcs(Reactions) + rcs(Balance) + rcs(ShotPower) + rcs(Jumping) + rcs(Stamina) + rcs(Strength) + rcs(LongShots) + rcs(Aggression) + rcs(Interceptions) + rcs(Positioning) + rcs(Vision) + rcs(Penalties) + rcs(Composure) + rcs(Marking) + rcs(StandingTackle) + rcs(SlidingTackle) + rcs(GKDiving) + rcs(GKHandling) + rcs(GKKicking) + rcs(GKPositioning) + rcs(GKReflexes), data = fifa19small)
```

# Create explainers

```{r createExplainers}
library("DALEX")
fifa_gbm_exp_deep <- explain(fifa_gbm_deep, 
                        data = fifa19small, 
                        y = fifa19small$SqrtValue^2, 
                        predict_function = function(m,x) 
                          predict(m, x, n.trees = 250)^2,
                        label = "GBM deep")

fifa_gbm_exp_shallow <- explain(fifa_gbm_shallow, 
                        data = fifa19small, 
                        y = fifa19small$SqrtValue^2, 
                        predict_function = function(m,x) 
                          predict(m, x, n.trees = 250)^2,
                        label = "GBM shallow")

fifa_rf_exp <- explain(fifa_rf, 
                        data = fifa19small, 
                        y = fifa19small$SqrtValue^2, 
                        predict_function = function(m,x) 
                          predict(m, x)$predictions^2,
                        label = "RF")

fifa_rms_exp <- explain(fifa_ols, 
                        data = fifa19small, 
                        y = fifa19small$SqrtValue^2, 
                        predict_function = function(m,x) 
                          predict(m, x)^2,
                        label = "RMS")
```

# Model performance

```{r modelPerforamnce}
library("auditor")

fifa_mr_gbm_shallow <- model_residual(fifa_gbm_exp_shallow)
fifa_mr_gbm_deep <- model_residual(fifa_gbm_exp_deep)
fifa_mr_gbm_rf <- model_residual(fifa_rf_exp)
fifa_mr_gbm_rms <- model_residual(fifa_rms_exp)

plot_residual_boxplot(fifa_mr_gbm_shallow, fifa_mr_gbm_deep, fifa_mr_gbm_rf, fifa_mr_gbm_rms) +
  scale_y_sqrt()

plot_prediction(fifa_mr_gbm_shallow, abline = TRUE) +
  scale_y_sqrt() +  scale_x_sqrt()
plot_prediction(fifa_mr_gbm_deep, abline = TRUE)  +
  scale_y_sqrt() +  scale_x_sqrt()
plot_prediction(fifa_mr_gbm_rf, abline = TRUE)  +
  scale_y_sqrt() +  scale_x_sqrt()
plot_prediction(fifa_mr_gbm_rms, abline = TRUE) +
  scale_y_sqrt() +  scale_x_sqrt()

```

# Feature importance

```{r featureImportance}
library("ingredients")
fifa_feat <- ingredients::feature_importance(fifa_gbm_exp_shallow)
plot(fifa_feat, max_vars = 12)

fifa_feat <- ingredients::feature_importance(fifa_gbm_exp_deep)
plot(fifa_feat, max_vars = 12)

fifa_feat <- ingredients::feature_importance(fifa_rf_exp)
plot(fifa_feat, max_vars = 12)

fifa_feat <- ingredients::feature_importance(fifa_rms_exp)
plot(fifa_feat, max_vars = 12)
```

# Partial Dependency Profiles

```{r pdp}
fifa19_pd_shallow <- ingredients::partial_dependency(fifa_gbm_exp_shallow, variables = c("Age", "Reactions","BallControl", "Dribbling"))

fifa19_pd_deep <- ingredients::partial_dependency(fifa_gbm_exp_deep, variables = c("Age", "Reactions","BallControl", "Dribbling"))

fifa19_pd_rf <- ingredients::partial_dependency(fifa_rf_exp, variables = c("Age", "Reactions","BallControl", "Dribbling"))

fifa19_pd_rms <- ingredients::partial_dependency(fifa_rms_exp, variables = c("Age", "Reactions","BallControl", "Dribbling"))

plot(fifa19_pd_shallow, fifa19_pd_deep, fifa19_pd_rf, fifa19_pd_rms) +
  scale_y_log10()
```

# Break Down

```{r breakDown}
library("iBreakDown")

fifa_pg <- break_down(fifa_gbm_exp_shallow, new_observation = fifa19small["R. Lewandowski",])
plot(fifa_pg)

fifa_pg <- break_down(fifa_gbm_exp_deep, new_observation = fifa19small["R. Lewandowski",])
plot(fifa_pg)

fifa_pg <- break_down(fifa_rf_exp, new_observation = fifa19small["R. Lewandowski",])
plot(fifa_pg)

fifa_pg <- break_down(fifa_rms_exp, new_observation = fifa19small["R. Lewandowski",])
plot(fifa_pg)

```

# Break Down with interactions

```{r iBreakDown}
library("iBreakDown")

fifa_pg <- break_down(fifa_gbm_exp_shallow, new_observation = fifa19small["R. Lewandowski",], interactions = TRUE)
plot(fifa_pg)

fifa_pg <- break_down(fifa_gbm_exp_deep, new_observation = fifa19small["R. Lewandowski",], interactions = TRUE)
plot(fifa_pg)

fifa_pg <- break_down(fifa_rf_exp, new_observation = fifa19small["R. Lewandowski",], interactions = TRUE)
plot(fifa_pg)

fifa_pg <- break_down(fifa_rms_exp, new_observation = fifa19small["R. Lewandowski",], interactions = TRUE)
plot(fifa_pg)

```

# Ceteris Paribus

```{r ceterisParibus}
fifa_cp_shallow <- ceteris_paribus(fifa_gbm_exp_shallow,
                           new_observation = fifa19small["R. Lewandowski",], variables = c("Age", "Reactions","BallControl", "Dribbling"),
                           variable_splits = list(Age = seq(15,45,0.1), Reactions = seq(20,100,0.1), BallControl = seq(20,100,0.1), Dribbling = seq(20,100,0.1))
                           )

fifa_cp_deep <- ceteris_paribus(fifa_gbm_exp_deep,
                           new_observation = fifa19small["R. Lewandowski",], variables = c("Age", "Reactions","BallControl", "Dribbling"),
                           variable_splits = list(Age = seq(15,45,0.1), Reactions = seq(20,100,0.1), BallControl = seq(20,100,0.1), Dribbling = seq(20,100,0.1))
                           )

fifa_cp_rf <- ceteris_paribus(fifa_rf_exp,
                           new_observation = fifa19small["R. Lewandowski",], variables = c("Age", "Reactions","BallControl", "Dribbling"),
                           variable_splits = list(Age = seq(15,45,0.1), Reactions = seq(20,100,0.1), BallControl = seq(20,100,0.1), Dribbling = seq(20,100,0.1))
                           )

fifa_cp_rms <- ceteris_paribus(fifa_rms_exp,
                           new_observation = fifa19small["R. Lewandowski",], variables = c("Age", "Reactions","BallControl", "Dribbling"),
                           variable_splits = list(Age = seq(15,45,0.1), Reactions = seq(20,100,0.1), BallControl = seq(20,100,0.1), Dribbling = seq(20,100,0.1))
                           )

plot(fifa_cp_shallow, fifa_cp_deep, fifa_cp_rf, fifa_cp_rms, color = "_label_") + 
  show_observations(fifa_cp_rf, fifa_cp_shallow, fifa_cp_deep,fifa_cp_rms, variables = c("Age", "Reactions","BallControl", "Dribbling")) + 
  scale_y_log10()

```

```

library(modelStudio)
fifa19_ms <- modelStudio(fifa_gbm_shallow, new_observation = fifa19small[c("Cristiano Ronaldo","R. Lewandowski"), ], B = 5, digits = 0)

op <- modelStudioOptions(
  margin.left = 6
)

print(fifa19_ms, options = op)

r2d3::save_d3_html(fifa19_ms, file = "fifa19.html")

```

