# Use Cases {-}

# FIFA 19 {#UseCaseFIFA}

In previous chapters we introduced a number of methods for instance level exploration of predictive models. In consecutive chapter we showed how to use Ceteris Paribus profiles, SHAP values, LIME or Break Down plots for models created on the dataset `titanic`. These examples we introduced and discussed separately as each of them was focused on a single method described in a given chapter.

In this chapter we present an example of full process for model development along the process introduces in Chapter \@ref(modelDevelopmentProcess). We will use a new dataset for FIFA 19 soccer game. Based on it we will tour through the process of data preparation, model assembly and model understanding. In each phase we show how to combine results from different methods of exploration.

The main goal of this chapter is to show how different techniques complement each other. Some phases, like data preparation, are simplified in order to leave space for the method for visual exploration and explanation of predictive models.





## Introduction

The story is following. The `https://sofifa.com/` portal is a reliable website for FIFA ratings of football players. Data from this website was scrapped and make available at the Kaggle webpage `https://www.kaggle.com/karangadiya/fifa19`.

We will use this data to build a predictive model for assessment of player value. Once the model will be created we will use methods for exploration and explanation to better understand how it is working and also to better understand which factors and how influence the player value.


## Data preparation 

The scraped data contains 89 columns, and various information about players along with photo, club, nationality and others. 
Here we will focus on 40 players statistics and the way how they influence model predictions.

The data set contains statistics for 16924 players. First, let's see distribution of selected variables from this dataset.

```{r warning=FALSE, message=FALSE, echo=FALSE}
set.seed(1313)
library("ggmosaic")
library("ggplot2")
library("DALEX")
library("patchwork")
library("scales")

euro_format <- function(largest_with_cents = 100000) {
  function(x) {
    x <- round(x, 0.01)
    if (max(x, na.rm = TRUE) < largest_with_cents &
        !all(x == floor(x), na.rm = TRUE)) {
      nsmall <- 2L
    } else {
      x <- round(x, 1)
      nsmall <- 0L
    }
    str_c("€", format(x, nsmall = nsmall, trim = TRUE, big.mark = ",", scientific = FALSE, digits=1L))
  }
}

load("misc/fifa19small.rda")
rownames(fifa19small) <- fifa19small$Name
```

Players values are heavily skewed. Half of players have estimated value between 0.3 and 2.2 millions of Euro. But few players have estimated values higher than 100 millions of Euro. Figure \@ref(fig:distFIFA19Value) presents empirical cumulative distribution function and histogram with log transformation of the OX axis.

```{r distFIFA19Value, warning=FALSE, message=FALSE, echo=FALSE, fig.width=9, fig.height=4.5, fig.cap="Empirical cumulative distribution function and histogram for values of players. The OX axis is in the log10 transformation.", out.width = '90%', fig.align='center'}
library("scales")

pl1 <- ggplot(fifa19small, aes(Value.EUR)) +
  stat_ecdf(geom = "step", pad = FALSE) +
  theme_drwhy() + 
  scale_x_continuous("Estimated value in Euro", trans = "log10", labels = dollar_format(suffix = "€", prefix = "")) + 
  ggtitle("ECDF for player's value") + 
  scale_y_continuous("Fraction of players with value higher than x", label = scales::percent)

pl2 <- ggplot(fifa19small, aes(Value.EUR)) +
  geom_histogram(bins = 50) +
  theme_drwhy() + 
  scale_x_continuous("Value in Euro", trans = "log10", labels = dollar_format(suffix = "€", prefix = "")) + 
  ylab("Number of players with given value") +
  ggtitle("Histogram for player's value") 

pl1 + pl2
```

Due to a large number of player characteristics we are not going to explore all of them but rather we will focus on four that will be discussed later in this chapter, namely: `Age`, `Reactions`, `BallControl` and `ShortPassing.`

Figure \@ref(fig:distFIFA19histograms) presents distributions for these variables. For `Age` we see that most players are between 20 and 30 years old. What is interesting in `BallControl` and `ShortPassing` is that they have bimodal distribution. The reason for that is that these characteristics are very low for goalkeepers but higher for other players. The variable `Reactions` has Gaussian shaped distribution with average 62 and standard deviation 9.

(ref:distFIFA19histogramsDesc) Histograms for selected characteristics of players. Note that `BallControl` and `ShortPassing` have bimodal distributions.

```{r distFIFA19histograms, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=6.5, fig.cap='(ref:distFIFA19histogramsDesc)', out.width = '90%', fig.align='center'}
fifa19small4 <- fifa19small[,c("Age", "Reactions", "BallControl", "ShortPassing")]

library("tidyr")
fifa19small4long <- gather(fifa19small4, variable, value)
ggplot(fifa19small4long, aes(value)) +
  geom_histogram() + 
  theme_drwhy() + facet_wrap(~variable, ncol = 2, scales = "free") + ggtitle("Histograms for player's characteristics") + scale_x_continuous("")
```

## Data understanding 

Time to see how these variables are linked with player's value.
Figure \@ref(fig:distFIFA19scatter) shows scatterplots for selected four characteristics. Because of the skewness of player's value the OY value is presented after log transformation.

For `Age` it looks like the relation is not monotonic, there is some optimal age in which players value is the highest, between 24 and 28 years. Value of youngest players are on average 10x lower. Same with olders players. 

For variables `BallControl` and `ShortPassing` the relation is not monotonic. In general the larger value of these coefficients the higher value of a player and most expensive are players with top characteristics. But among players with very low scores in `BallControl` and `ShortPassing` some are very expensive too. As it was suggested earlier, these players are probably goalkeepers.

For variable `Reactions` the link with player's value is monotonic. As expected, the higher `Reactions` the higher player's value.


```{r distFIFA19scatter, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=6.5, fig.cap="Scatterplot for relation between selected four players characteristics and values of players.", out.width = '90%', fig.align='center'}
fifa19small4v <- fifa19small[,c("Value.EUR","Age", "Reactions", "BallControl", "ShortPassing")]

fifa19small4long <- gather(fifa19small4v, variable, value, -Value.EUR)

ggplot(fifa19small4long, aes(value, Value.EUR)) +
  geom_point() + geom_smooth(size = 2, se = FALSE) +
  theme_drwhy() + 
  facet_wrap(~variable, ncol = 2, scales = "free") + 
  scale_y_continuous("Value in Euro", trans = "log10", labels = dollar_format(suffix = "€", prefix = "")) + 
  scale_x_continuous("") + 
  ggtitle("Scatterplots for player's characteristics")

```

Figure \@ref(fig:distFIFA19scatter2) shows pairwise scatterplots for dependent variables. Three observations are clear from these scatterplots. One is that `Age` has positive correlation with other variables. On average older players have higher skills. Second is that skills are positively correlated, the correlation between `BallControl` and `ShortPassing` is higher than 0.9. Third is that goalkeepers' characteristics are different than rest of players.


```{r distFIFA19scatter2, warning=FALSE, message=FALSE, echo=FALSE, fig.width=10, fig.height=9, fig.cap="Scatterplot for relation between selected four players characteristics and values of players.", out.width = '90%', fig.align='center'}
library("GGally")
ggpairs(fifa19small4v[,-1], 
        diag = list(continuous = "barDiag")) +
  theme_drwhy() + 
   ggtitle("Scatterplot matrix for player's characteristics")

```

Let's compare results from this data exploration with exploration of predictive models that will be fitted on this data.


## Model assembly 

Time to build a predictive model for players' value based on selected characteristics. We will use a trained elastic model to explore the relation between players' characteristics and players' values.

Having clean data then model assembly is easy. For FIFA 19 data we will try four models with different structures that are able to catch different types of relations. One model would be enough, but we will try four different models to see if they catch similar relations.

Considered models are:

- boosting model with 250 trees 1 level depth as implemented in package `gbm` [@gbm],
- boosting model with 250 trees 4 levels depth, this model shall be able to catch interactions between features,
- linear model with spline transformation of dependent variables implemented in package `rms` [@rms],
- random forest model with 250 trees as implemented in package `ranger` [@rangerRpackage].

```{r, warning=FALSE, message=FALSE, echo=FALSE}
# log10 transfromation
fifa19small <- fifa19small[fifa19small$Value.EUR > 1, ]
fifa19small$LogValue <- log10(fifa19small$Value.EUR)
fifa19small <- fifa19small[,-c(1, 2, 3, 4, 6)]
```

```{r createModels, warning=FALSE, message=FALSE}
library("gbm")
fifa_gbm_shallow <- gbm(LogValue~., data = fifa19small, n.trees = 250,
        interaction.depth = 1, distribution = "gaussian")

fifa_gbm_deep <- gbm(LogValue~., data = fifa19small, n.trees = 250,
        interaction.depth = 4, distribution = "gaussian")

library("ranger")
fifa_rf <- ranger(LogValue~., data = fifa19small, num.trees = 250)

library("rms")
fifa_ols <- ols(LogValue ~ rcs(Age) + rcs(International.Reputation) +
        rcs(Skill.Moves) + rcs(Crossing) + rcs(Finishing) +
        rcs(HeadingAccuracy) + rcs(ShortPassing) + rcs(Volleys) +
        rcs(Dribbling) + rcs(Curve) + rcs(FKAccuracy) +
        rcs(LongPassing) + rcs(BallControl) + rcs(Acceleration) +
        rcs(SprintSpeed) + rcs(Agility) + rcs(Reactions) +
        rcs(Balance) + rcs(ShotPower) + rcs(Jumping) + rcs(Stamina) +
        rcs(Strength) + rcs(LongShots) + rcs(Aggression) +
        rcs(Interceptions) + rcs(Positioning) + rcs(Vision) +
        rcs(Penalties) + rcs(Composure) + rcs(Marking) +
        rcs(StandingTackle) + rcs(SlidingTackle) + rcs(GKDiving) +
        rcs(GKHandling) + rcs(GKKicking) + rcs(GKPositioning) +
        rcs(GKReflexes), data = fifa19small)
```

Before we can explore model behavior we need to create explainers with the `DALEX::explain` function. These explainers will be later used to assess model performance.

Note that models are trained on logarithm of the value, but it will be much more natural to operate on values in Euro. This is why in explainers we specified a user defined predict function that transforms log value to the value in Euro.

Each explainer got also a unique `label` and corresponding `data` and `y` arguments.

```{r createExplainers, message=FALSE, warning=FALSE, results='hide'}
library("DALEX")
fifa_gbm_exp_deep <- explain(fifa_gbm_deep, 
        data = fifa19small, y = 10^fifa19small$LogValue, 
        predict_function = function(m,x) 10^predict(m, x, n.trees = 250),
        label = "GBM deep")

fifa_gbm_exp_shallow <- explain(fifa_gbm_shallow, 
        data = fifa19small, y = 10^fifa19small$LogValue, 
        predict_function = function(m,x) 10^predict(m, x, n.trees = 250),
        label = "GBM shallow")

fifa_rf_exp <- explain(fifa_rf, 
        data = fifa19small, y = 10^fifa19small$LogValue, 
        predict_function = function(m,x) 10^predict(m, x)$predictions,
        label = "RF")

fifa_rms_exp <- explain(fifa_ols, 
        data = fifa19small, y = 10^fifa19small$LogValue, 
        predict_function = function(m,x) 10^predict(m, x),
        label = "RMS")
```

## Model audit

We have created four models. Let's see which model is better. Figure \@ref(fig:modelPerforamanceBoxplot) compares distributions of absolute model residuals. Crosses corresponds to average, which correspond to RMSE. On average, smallest residuals are for the Random Forest model.

```{r modelPerforamanceBoxplot, warning=FALSE, message=FALSE, echo=TRUE, fig.width=8, fig.height=3, fig.cap="Distribution of absolute values of residuals. The smaller are values the better is the model. Dots stand for averages.", out.width = '90%', fig.align='center'}
library("DALEX")
(fifa_mr_gbm_shallow <- model_performance(fifa_gbm_exp_shallow))
(fifa_mr_gbm_deep <- model_performance(fifa_gbm_exp_deep))
(fifa_mr_gbm_rf <- model_performance(fifa_rf_exp))
(fifa_mr_gbm_rms <- model_performance(fifa_rms_exp))

plot(fifa_mr_gbm_shallow, fifa_mr_gbm_deep, fifa_mr_gbm_rf, fifa_mr_gbm_rms, geom = "boxplot") +
  scale_y_continuous("Absolute residuals in Euro", trans = "log10", labels = dollar_format(suffix = "€", prefix = "")) + 
  ggtitle("Distributions of model absolute residuals")
```

But performance is not everything. Figure \@ref(fig:modelPerformanceScatterplot) show diagnostics plots for every model. Each scatterplot shows true target variable against model predictions. The random forest model has predictions closest to the true target values. 

Extreme predictions (lowest and highest) are biased towards the mean, what is typical for such type of models. This means that Random Forest models learned factors that influence players' values, but for the most expensive players these values will be underestimated.


```{r modelPerformanceScatterplot, warning=FALSE, message=FALSE, echo=TRUE, fig.width=10, fig.height=10, fig.cap="Diagnostics plots Predicted vs. True target values. Points correspond to particular players. The closer to the diagonal the better is the model.", out.width = '90%', fig.align='center'}
fifa_md_gbm_shallow <- model_diagnostics(fifa_gbm_exp_shallow)
fifa_md_gbm_deep <- model_diagnostics(fifa_gbm_exp_deep)
fifa_md_gbm_rf <- model_diagnostics(fifa_rf_exp)
fifa_md_gbm_rms <- model_diagnostics(fifa_rms_exp)

plot(fifa_md_gbm_shallow, fifa_md_gbm_deep, 
                fifa_md_gbm_rf, fifa_md_gbm_rms, 
     variable = "y", yvariable = "y_hat") +
  scale_x_continuous("Value in Euro", trans = "log10", labels = dollar_format(suffix = "€", prefix = "")) + 
  scale_y_continuous("Estimated value in Euro", trans = "log10", labels = dollar_format(suffix = "€", prefix = "")) + 
  facet_wrap(~label) +
  geom_abline(slope = 1) + theme(legend.position = "none") +
  ggtitle("Diagnostics plot Predicted vs True target values", "")
```



## Model understanding

Figure \@ref(fig:featureImportance) shows variable importance plots for four selected models. Only 12 most important variables in each model are presented.

Some variables are important for all models, like `Reactions` or `BallControl`. Importance of other variables may be very different. All models except random forest are using some characteristics of goalkeepers.


```{r featureImportance, warning=FALSE, message=FALSE, echo=TRUE, fig.width=9, fig.height=11, fig.cap="Variable importance plots for four considered models. Each bar starts in a RMSE for the model and ends in a RMSE calculated for data with permuted single variable.", out.width = '90%', fig.align='center'}
fifa_mp_gbm_shallow <- model_parts(fifa_gbm_exp_shallow)
fifa_mp_gbm_deep <- model_parts(fifa_gbm_exp_deep)
fifa_mp_rf <- model_parts(fifa_rf_exp)
fifa_mp_rms <- model_parts(fifa_rms_exp)

plot(fifa_mp_gbm_shallow, fifa_mp_gbm_deep, 
     fifa_mp_rf, fifa_mp_rms,
     max_vars = 20, bar_width = 4, show_boxplots = FALSE) 
```

Figure \@ref(fig:usecaseFIFApdp) shows Partial Dependence profiles for the most important variables. They show average relation between particular variable and players value.

The general direction of relation in all models is the same. The larger the player characteristic the higher is the price. With a single exception – variable Age.

Random forest model has smallest range of average model responses. All tree-based models stabilize average predictions at the ends of variables ranges.


The most interesting difference between Exploratory Data Analysis presented in Figure \@ref(fig:distFIFA19scatter) and Exploratory Model Analysis presented in Figure \@ref(fig:usecaseFIFApdp) is related with variable `Age`. In Figure \@ref(fig:distFIFA19scatter) the relation was non-monotonic while in Figure \@ref(fig:usecaseFIFApdp) its monotonically decreasing. 
How we can explain this difference? 
One explanation is following: Youngest players have lower values not because of their age but because of lower skills that are correlated with Age. The EDA analysis cannot entangle these effects, thus for youngest players we see lover values also because their lower skills. But models learned that once we take skills into account, the effect of age is only decreasing. 

This example also shows, that proper *exploration of models may be more insightful than exploration of raw data*. Variable `Age` is correlated with other confounding variables. This entangle was visible in the EDA analysis. But models learned to disentangle these effects.


```{r usecaseFIFApdp, warning=FALSE, message=FALSE, echo=TRUE, fig.width=9, fig.height=8, fig.cap="Partial Dependence profiles for four selected variables and four considered models.", out.width = '90%', fig.align='center'}
selected_variables <- c("Age", "Reactions","BallControl", "ShortPassing")

fifa19_pd_shallow <- model_profile(fifa_gbm_exp_shallow, variables = selected_variables)$agr_profiles
fifa19_pd_deep <- model_profile(fifa_gbm_exp_deep, variables = selected_variables)$agr_profiles
fifa19_pd_rf <- model_profile(fifa_rf_exp, variables = selected_variables)$agr_profiles
fifa19_pd_rms <- model_profile(fifa_rms_exp, variables = selected_variables)$agr_profiles

plot(fifa19_pd_shallow, fifa19_pd_deep, fifa19_pd_rf, fifa19_pd_rms) +
  scale_y_continuous("Estimated value in Euro", trans = "log10", labels = dollar_format(suffix = "€", prefix = "")) + 
  ggtitle("Partial Dependence profiles for selected variables")
```


## Instance understanding

Time to see how the model behaves for a single observation / player 
This can be done for any player, but for this example we will use *Robert Lewandowski*, the most valuable polish football player.

Here are his characteristics in the FIFA 19 database.

```{r RobertLewandowski}
fifa19small["R. Lewandowski",]
```

In the chapter \@ref(breakDown) we showed a Break Down plots for presentation of variable attributions. In the Figure \@ref(fig:usecaseFIFAbreakDown) we show Break Down plots for Robert Lewandowski predictions.

```{r usecaseFIFAbreakDown, warning=FALSE, message=FALSE, echo=FALSE, fig.width=12, fig.height=6, fig.cap="Break down plot for Robert Lewandowski. Results for GBM and RF model.", out.width = '100%', fig.align='center'}
fifa_bd_gbm <- predict_parts(fifa_gbm_exp_shallow, new_observation = fifa19small["R. Lewandowski",])
fifa_bd_rf <- predict_parts(fifa_rf_exp, new_observation = fifa19small["R. Lewandowski",])

pl1 <- plot(fifa_bd_gbm) +
  scale_y_continuous("Estimated value in Euro", labels = dollar_format(suffix = "€", prefix = "")) + 
  ggtitle("Break Down plot for Robert Lewandowski")

pl2 <- plot(fifa_bd_rf) +
  scale_y_continuous("Estimated value in Euro", labels = dollar_format(suffix = "€", prefix = "")) + 
  ggtitle("Break Down plot for Robert Lewandowski")

pl1 + pl2
```

In the chapter \@ref(shapley) we showed a SHAP values for presentation of variable attributions. In the Figure \@ref(fig:usecaseFIFAshap) we show SHAP plots for Robert Lewandowski predictions. As it was expected, these explanations are consistent.

```{r usecaseFIFAshap, warning=FALSE, message=FALSE, echo=FALSE, fig.width=7, fig.height=5.5, fig.cap="SHAP values for GBM model.", out.width = '60%', fig.align='center'}
fifa_pg <- predict_parts(fifa_gbm_exp_shallow, new_observation = fifa19small["R. Lewandowski",],
                         type = "shap")
plot(fifa_pg, show_boxplots = FALSE) +
  scale_y_continuous("Estimated value in Euro", labels = dollar_format(suffix = "€", prefix = "")) + 
  ggtitle("SHAP values plot for Robert Lewandowski (GBM model)")
```


Robert Lewandowski is a striker. It makes sense that his most valuable characteristics are `Reactions` and `BallControl.`

How these plots will look like for goalkeepers? Figure \@ref(fig:usecaseFIFAbreakDownWS) show Break Down plots for Wojciech Szczęsny - most valuable polish goalkeeper. As we see the most important coefficients make sense, most of them are linked with properties of goalkeepers.


```{r usecaseFIFAbreakDownWS, warning=FALSE, message=FALSE, echo=FALSE, fig.width=12, fig.height=6, fig.cap="Break down plot for Wojciech Szczęsny. Results for GBM and RF model.", out.width = '100%', fig.align='center'}
library("iBreakDown")
fifa_bd_gbm <- predict_parts(fifa_gbm_exp_shallow, new_observation = fifa19small["W. Szczęsny",])
fifa_bd_rf <- predict_parts(fifa_rf_exp, new_observation = fifa19small["W. Szczęsny",])

pl1 <- plot(fifa_bd_gbm) +
  scale_y_continuous("Estimated value in Euro", labels = dollar_format(suffix = "€", prefix = "")) + 
  ggtitle("Break Down plot for Wojciech Szczęsny (GBM model)")

pl2 <- plot(fifa_bd_rf) +
  scale_y_continuous("Estimated value in Euro", labels = dollar_format(suffix = "€", prefix = "")) + 
  ggtitle("Break Down plot for Wojciech Szczęsny (RF model)")

pl1 + pl2
```


In Chapter \@ref(ceterisParibus) we introduced Ceteris Paribus profiles. These are more details steps of the model exploration. Based on an example of Robert Lewandowski. Figure \@ref(fig:usecaseFIFAceterisParibus) show how change in one characteristic affects model value.

Tree based models are flat on borders, and for them Robert value is the highest for variables `Reactions`, `BallControl` or `Dribling`. When it comes to Age we see that the predicted value is just before a larger drop in value prediction.


```{r usecaseFIFAceterisParibus, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=6.5, fig.cap="Ceteris Paribus profiles for Robert Lewandowski for four selected variables.", out.width = '90%', fig.align='center'}
selected_splits <- list(Age = seq(15,45,0.1), Reactions = seq(20,100,0.1), BallControl = seq(20,100,0.1), ShortPassing = seq(20,100,0.1))

fifa_cp_shallow <- predict_profile(fifa_gbm_exp_shallow,
                           new_observation = fifa19small["R. Lewandowski",], variables = selected_variables,
                           variable_splits = selected_splits)

fifa_cp_deep <- predict_profile(fifa_gbm_exp_deep,
                           new_observation = fifa19small["R. Lewandowski",], variables = selected_variables,
                           variable_splits = selected_splits)

fifa_cp_rf <- predict_profile(fifa_rf_exp,
                           new_observation = fifa19small["R. Lewandowski",], variables = selected_variables,
                           variable_splits = selected_splits)

fifa_cp_rms <- predict_profile(fifa_rms_exp,
                           new_observation = fifa19small["R. Lewandowski",], variables = selected_variables,
                           variable_splits = selected_splits)

plot(fifa_cp_shallow, fifa_cp_deep, fifa_cp_rf, fifa_cp_rms, color = "_label_", variables = c("Age", "Reactions", "BallControl", "ShortPassing")) + 
  scale_y_continuous("Estimated value in Euro", trans = "log10", labels = dollar_format(suffix = "€", prefix = "")) 

```

Figure \@ref(fig:usecaseFIFAceterisParibusNeighbours) shows residuals for all observations against residuals for 50 closest neighbours of Robert Lewandowski. Clearly, among neighbours he has the most expensive players and therefore their residuals are much higher than average residuals. This was also visible in the right top corner of each panel in Figure \@ref(fig:modelPerformanceScatterplot).


```{r usecaseFIFAceterisParibusNeighbours, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=4.5, fig.cap="Distribution of residuals for all players and neighbours of Robert Lewandowski.", out.width = '90%', fig.align='center'}
id_rf_v6 <- individual_diagnostics(fifa_gbm_exp_shallow,
                          fifa19small["R. Lewandowski",],
                          neighbours = 50)

plot(id_rf_v6) 
```

Figure \@ref(fig:usecaseFIFAceterisParibusNeighboursAgeGBM) shows local-fidelity plot for 15 neighbours of Robert Lewandowski. Model behaviour for these neighbours are similar. We also see that the most expensive players are undervalued by the model.

```{r usecaseFIFAceterisParibusNeighboursAgeGBM, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=7.5, fig.cap="Ceteris Paribus profiles for 15 neighbours of Robert Lewandowski.", out.width = '90%', fig.align='center'}
id_rf_v6 <- individual_diagnostics(fifa_gbm_exp_shallow,
                          fifa19small["R. Lewandowski",],
                          neighbours = 15,
                       variables = "Age")
plot(id_rf_v6) +  
  scale_y_continuous("Estimated value in Euro", trans = "log10", labels = dollar_format(suffix = "€", prefix = "")) + 
  ggtitle("Local-fidelity plot for Robert", "Calculated for the fifa_gbm_exp_shallow model")
```

## CR7 

In this section we present model explanations for Cristiano Ronaldo (CR7). Here are his characteristics in the FIFA 19 database.

```{r CR7}
fifa19small["Cristiano Ronaldo",]
```

Let's start with Break Down plots for variable attributions. 

```{r usecaseFIFAbreakDownCR7, warning=FALSE, message=FALSE, echo=FALSE, fig.width=12, fig.height=6, fig.cap="Break down plot for Cristiano Ronaldo.", out.width = '100%', fig.align='center'}
fifa_bd_gbm <- predict_parts(fifa_gbm_exp_shallow, 
        new_observation = fifa19small["Cristiano Ronaldo",])

plot(fifa_bd_gbm) +
  scale_y_continuous("Estimated value in Euro", labels = dollar_format(suffix = "€", prefix = "")) + 
  ggtitle("Break Down plot for Cristiano Ronaldo")
```

Cristiano Ronaldo is a striker. It makes sense that his most valuable characteristics are `Reactions` and `BallControl.` Let's see Ceteris Paribus profiles for him. 

```{r usecaseFIFAceterisParibusCR7, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=6.5, fig.cap="Ceteris Paribus profiles for Cristiano Ronaldo for four selected variables.", out.width = '90%', fig.align='center'}
selected_splits <- list(Age = seq(15,45,0.1), Reactions = seq(20,100,0.1), BallControl = seq(20,100,0.1), ShortPassing = seq(20,100,0.1))

fifa_cp_shallow <- predict_profile(fifa_gbm_exp_shallow,
                           new_observation = fifa19small["Cristiano Ronaldo",], variables = selected_variables,
                           variable_splits = selected_splits)
plot(fifa_cp_shallow, variables = c("Age", "Reactions", "BallControl", "ShortPassing")) + 
  scale_y_continuous("Estimated value in Euro", trans = "log10", labels = dollar_format(suffix = "€", prefix = "")) 

```











```{r warning=FALSE, message=FALSE, echo=FALSE, eval=FALSE}
# examples for DALEX blogpost

set.seed(1313)
library("ggplot2")
library("DALEX")

load("misc/fifa19small.rda")
rownames(fifa19small) <- fifa19small$Name
# log10 transfromation
fifa19small <- fifa19small[fifa19small$Value.EUR > 1, ]
fifa19small$LogValue <- log10(fifa19small$Value.EUR)
fifa19small <- fifa19small[,-c(1, 2, 3, 4, 6)]

library("gbm")
fifa_gbm_shallow <- gbm(LogValue~., data = fifa19small, n.trees = 250, interaction.depth = 1, distribution = "gaussian")

fifa_gbm <- gbm(LogValue~., 
                data = fifa19small, 
                n.trees = 250, 
                interaction.depth = 4, 
                distribution = "gaussian")

library("DALEX")
fifa_exp <- explain(fifa_gbm, 
                data = fifa19small, 
                y = 10^fifa19small$LogValue, 
                predict_function = function(m,x) 10^predict(m, x, n.trees = 250),
                label = "GBM deep")

cr17 <- fifa19small["Cristiano Ronaldo",]

predict(fifa_exp, cr17)


euro_format <- function(largest_with_cents = 100000) {
  function(x) {
    x <- round(x, 0.01)
    if (max(x, na.rm = TRUE) < largest_with_cents &
        !all(x == floor(x), na.rm = TRUE)) {
      nsmall <- 2L
    } else {
      x <- round(x, 1)
      nsmall <- 0L
    }
    str_c("€", format(x, nsmall = nsmall, trim = TRUE, big.mark = ",", scientific = FALSE, digits=1L))
  }
}


library(dplyr)
library(scales)
predict_parts(fifa_exp, cr17) %>% plot()
predict_parts(fifa_exp, cr17) %>% plot() + ggtitle("Break Down for CR7") + scale_y_continuous("",labels = dollar_format(suffix = "€", prefix = ""), limits=c(0,55000000))


predict_parts(fifa_exp, cr17, type = "shap") %>% plot(show_boxplots = FALSE) -> a
a + ggtitle("Shapley for CR7") + scale_y_continuous("",labels = dollar_format(suffix = "€", prefix = ""), limits=c(-16000000,16000000))

predict_profile(fifa_exp, cr17) %>% plot(variables = c("Age", "BallControl")) + ggtitle("Ceteris Paribus Profiles for CR7") + scale_y_continuous("",labels = dollar_format(suffix = "€", prefix = ""))

plot(a, show_boxplots = FALSE) + coord_flip(xlim=c(31,41)) + ggtitle("Shapley for CR17")

predict_diagnostics(fifa_exp, cr17) %>% plot() + scale_x_continuous("",labels = dollar_format(suffix = "€", prefix = ""))

model_performance(fifa_exp) %>% plot()

model_parts(fifa_exp) %>% plot(show_boxplots = FALSE, max_vars = 10)

model_profile(fifa_exp) %>% plot(variables = c("Age", "BallControl"), geom = "profiles") + ggtitle("Partial Dependence Profiles for GBM model")

model_profile(fifa_exp) %>% DALEX:::plot.model_profile_profiles(variables = c("Age", "BallControl")) + scale_y_continuous("",labels = dollar_format(suffix = "€", prefix = ""))+ ggtitle("Partial Dependence Profiles for GBM model")

```

