# Use Case for FIFA 19 {#UseCaseFIFA}

In previous chapters we introduced a number of methods for instance level exploration of predictive models. In consecutive chapter we showed how to use Ceteris Poribus profiles, SHAP values, LIME or Break Down plots for models created on the dataset `titanic`. These examples we introduced and discussed separately as each of them was focused on a single method described in a given chapter.

In this chapter we present an example of full process for model development along the process introduces in chapter \@ref(modelDevelopmentProcess). We will use a new dataset for FIFA 19 soccer game. Based on it we will tour through the process of data preparation, model assembly and model understanding. In each phase we show how to combine results from different methods of exploration.

The main goal of this chapter is to show how different techniques complement each other. Some phases, like data preparation, are simplified in order to leave space for the method for visual exploration and explanation of predictive models.





## Introduction

The story is following. The `https://sofifa.com/` portal is a reliable website for FIFA ratings of football players. Data from this website was scrapped and make available at the Kaggle webpage `https://www.kaggle.com/karangadiya/fifa19`.

We will use this data to build a predictive model for assessment of player value. Once the model will be created we will use methods for exploration and explanation to better understand how it is working and also to better understand which factors and how influence the player value.


## Data preparation 

The scrapped data contains 89 columns, and variuos information about players along with photo, club, nationality and others. 
Here we will focus on 40 players statistics and the way how they influence model predictions.

The data set contains statistics for 16924 players. First, let's see distribution of selected variables from this dataset.

```{r warning=FALSE, message=FALSE, echo=FALSE}
set.seed(1313)
library("ggmosaic")
library("ggplot2")
library("DALEX")

load("misc/fifa19small.rda")
rownames(fifa19small) <- fifa19small$Name
```

Player value is heavily skewed. Half of players have estimated values between 0.3 and 2.2 millions of Euro. But few players have estimated values higher than 100 millions of Euro. Below we present the empirical cumulative distribution function with log transformation of the OX axis.

```{r distFIFA19Value, warning=FALSE, message=FALSE, echo=FALSE, fig.width=5, fig.height=4.5, fig.cap="(fig:distFIFA19Value) Empirical cummulative distribution function for values of players. The OX axis is in the log10 transformation.", out.width = '60%', fig.align='center'}
ggplot(fifa19small, aes(Value.EUR)) +
  stat_ecdf(geom = "step", pad = FALSE) +
  theme_drwhy() + scale_x_log10(name = "Estimated value in Euro") + ylab("Fraction of players with value higher than x") +
  ggtitle("ECDF for player's value") + scale_y_continuous(label = scales::percent)
```

Due to a large number of player characteristics we are not going to explore all of them but rather we will focus on four that will be discussed later in this chapter, namely: Age, Reactions, BallControl and ShortPassing.

Figure \@ref(fig:distFIFA19histograms) presents distributions for these variables. For `Age` we see that most players are between 20 and 30 years old. What is interesting in `BallControl` and `ShortPassing` is that they have bimodal distribution. The reason for that is that these characteristics are very low for goalkeepers but higher for other players. The variable `Reactions` has Gaussian shaped distributtion with average 62 and standard deviation 9.

```{r distFIFA19histograms, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=6.5, fig.cap="(fig:distFIFA19histograms) Distribution of selected characteristics of players.", out.width = '90%', fig.align='center'}
fifa19small4 <- fifa19small[,c("Age", "Reactions", "BallControl", "ShortPassing")]

library("tidyr")
fifa19small4long <- gather(fifa19small4, variable, value)
ggplot(fifa19small4long, aes(value)) +
  geom_histogram() + 
  theme_drwhy() + facet_wrap(~variable, ncol = 2, scales = "free") + ggtitle("Histograms for player's characteristics") + scale_x_continuous("")
```

## Data understanding 

Time to see how these variables are linked with player's value.
Figure \@ref(fig:distFIFA19scatter) shows scatterplots for selected four characteristics. Because of the skewness of player's value the OY value is presented after log transformation.

For `Age` it looks like the relation is not monotonic, there is some optimal age in which players value is the highest, between 24 and 28 years. Value of youngest players are on average 10x lower, same with olders players. 

For variables `BallControl` and `ShortPassing` the relation is not monotonic. In general the larger value of these coefficients the higher value of a player and most expensive are players with top characteristics. But among players with very low scores in `BallControl` and `ShortPassing` some are very expensive too. As it was suggested earlier, these players are probably goalkeepers.

For variable `Reactions` the link with player's value is monotonic. As expected, the higher `Reactions` the higher player's value.


```{r distFIFA19scatter, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=6.5, fig.cap="(fig:distFIFA19scatter) Scatterplot for realtion between selected four players characteristics and values of players.", out.width = '90%', fig.align='center'}
fifa19small4v <- fifa19small[,c("Value.EUR","Age", "Reactions", "BallControl", "ShortPassing")]

fifa19small4long <- gather(fifa19small4v, variable, value, -Value.EUR)

ggplot(fifa19small4long, aes(value, Value.EUR)) +
  geom_point() + geom_smooth(size = 2, se = FALSE) +
  theme_drwhy() + facet_wrap(~variable, ncol = 2, scales = "free") + scale_y_log10() + scale_x_continuous("")+ ggtitle("Scatterplots for player's characteristics")

```


Let's compare results from this data exploration with exploration of predictive models that will be fitted on this data.


## Model assembly 

Time to build a predictive model for player's value based on selected characteristics. We will quickly train some elastic model to further explore the relation between player's characteristics and value.

If the data is clean then building a model is easy. For FIFA 19 data we will try four models with different structures that are able to catch different types of relations. Considered models are:

- boosting model with 250 trees 1 level depth as implemented in package `gbm` [@gbm],
- boosting model with 250 trees 4 levels depth, this model shall be able to catch interactions between features,
- linear model with spline transformation of dependent variables implemented in package `rms` [@rms],
- random forest model with 250 trees as implemented in package `ranger` [@rangerRpackage].

```{r, warning=FALSE, message=FALSE, echo=FALSE}
# log10 transfromation
fifa19small <- fifa19small[fifa19small$Value.EUR > 1, ]
fifa19small$LogValue <- log10(fifa19small$Value.EUR)
fifa19small <- fifa19small[,-c(1, 2, 3, 4, 6)]
```

```{r createModels, warning=FALSE, message=FALSE}
library("gbm")
fifa_gbm_shallow <- gbm(LogValue~., data = fifa19small, n.trees = 250, interaction.depth = 1, distribution = "gaussian")

fifa_gbm_deep <- gbm(LogValue~., data = fifa19small, n.trees = 250, interaction.depth = 4, distribution = "gaussian")

library("ranger")
fifa_rf <- ranger(LogValue~., data = fifa19small, num.trees = 250)

library("rms")
fifa_ols <- ols(LogValue ~ rcs(Age) + rcs(International.Reputation) + rcs(Skill.Moves) + rcs(Crossing) + rcs(Finishing) + rcs(HeadingAccuracy) + rcs(ShortPassing) + rcs(Volleys) + rcs(Dribbling) + rcs(Curve) + rcs(FKAccuracy) + rcs(LongPassing) + rcs(BallControl) + rcs(Acceleration) + rcs(SprintSpeed) + rcs(Agility) + rcs(Reactions) + rcs(Balance) + rcs(ShotPower) + rcs(Jumping) + rcs(Stamina) + rcs(Strength) + rcs(LongShots) + rcs(Aggression) + rcs(Interceptions) + rcs(Positioning) + rcs(Vision) + rcs(Penalties) + rcs(Composure) + rcs(Marking) + rcs(StandingTackle) + rcs(SlidingTackle) + rcs(GKDiving) + rcs(GKHandling) + rcs(GKKicking) + rcs(GKPositioning) + rcs(GKReflexes), data = fifa19small)
```

Before we can explore model behavior we need to create explainers with the `DALEX::explain` function. These explainers will be later used to asses model performance.

Note that models are trained on logarithm of the value, but it will be much more natural to operate on values in Euro. This is why in explainers we specified a user defined predict function that transforms log value to the value in Euro.

Each explainer got also a unique `label` and corresponding `data` and `y` arguments.

```{r createExplainers, message=FALSE, warning=FALSE, results='hide'}
library("DALEX")
fifa_gbm_exp_deep <- explain(fifa_gbm_deep, 
                        data = fifa19small, y = 10^fifa19small$LogValue, 
                        predict_function = function(m,x) 10^predict(m, x, n.trees = 250),
                        label = "GBM deep")

fifa_gbm_exp_shallow <- explain(fifa_gbm_shallow, 
                        data = fifa19small, y = 10^fifa19small$LogValue, 
                        predict_function = function(m,x) 10^predict(m, x, n.trees = 250),
                        label = "GBM shallow")

fifa_rf_exp <- explain(fifa_rf, 
                        data = fifa19small, y = 10^fifa19small$LogValue, 
                        predict_function = function(m,x) 10^predict(m, x)$predictions,
                        label = "RF")

fifa_rms_exp <- explain(fifa_ols, 
                        data = fifa19small, y = 10^fifa19small$LogValue, 
                        predict_function = function(m,x) 10^predict(m, x),
                        label = "RMS")
```

## Model audit

We have createwd four models. Let's see which model is better. Figure \@ref(fig:modelPerforamanceBoxplot) compares absolute model residuals. On average, smallest residuals are for the Random Forest model.

```{r modelPerforamanceBoxplot, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=4, fig.cap="(fig:modelPerforamanceBoxplot) Distribution of absolute values of residuals. The smaller are values the better is the model. Crosses stand for averages.", out.width = '70%', fig.align='center'}
library("auditor")

fifa_mr_gbm_shallow <- model_residual(fifa_gbm_exp_shallow)
fifa_mr_gbm_deep <- model_residual(fifa_gbm_exp_deep)
fifa_mr_gbm_rf <- model_residual(fifa_rf_exp)
fifa_mr_gbm_rms <- model_residual(fifa_rms_exp)

plot_residual_boxplot(fifa_mr_gbm_shallow, fifa_mr_gbm_deep, fifa_mr_gbm_rf, fifa_mr_gbm_rms) +
  scale_y_log10() +
  ggtitle("Distributions of model residuals")
```

But performance is not everything. Figure \@ref(fig:modelPerformanceScatterplot) show diagnostic plots for every model. Each scatterplot shows true target variable against model predictions. The random forest model has predictions closest to the true target values. Extreme predictions (lowest and highest) are biased towards the mean, what is typical for such type of models.


```{r modelPerformanceScatterplot, warning=FALSE, message=FALSE, echo=FALSE, fig.width=9, fig.height=9, fig.cap="(fig:modelPerformanceScatterplot) Diagnostic plots Predicted vs. True target values. Points correspond to particular players. The closer to the diagonal the better is the model.", out.width = '100%', fig.align='center'}
# plot_prediction(fifa_mr_gbm_shallow, abline = TRUE) +
#   scale_y_log10() +  scale_x_log10()
# plot_prediction(fifa_mr_gbm_deep, abline = TRUE)  +
#   scale_y_log10() +  scale_x_log10()
# plot_prediction(fifa_mr_gbm_rf, abline = TRUE)  +
#   scale_y_log10() +  scale_x_log10()
# plot_prediction(fifa_mr_gbm_rms, abline = TRUE) +
#   scale_y_log10() +  scale_x_log10()

plot_prediction(fifa_mr_gbm_shallow, fifa_mr_gbm_deep, 
                fifa_mr_gbm_rf, fifa_mr_gbm_rms, abline = TRUE) +
  scale_y_log10() +  scale_x_log10() +
  facet_wrap(~`_label_`) + theme(legend.position = "none") +
  ggtitle("Diagnostic plot Predicted vs True target values")

```



## Model understanding

Figure \@ref(fig:featureImportance) shows variable importance plots for four selected models. Only 12 most important variables in each model are presented.

Some variables are important for all models, like `Reactions` or `BallControl`. Importance of other variables may be very different. All models except random forest are using some characteristics of goalkeepers.


```{r featureImportance, warning=FALSE, message=FALSE, echo=FALSE, fig.width=10, fig.height=7, fig.cap="(fig:featureImportance) Variable importance plots for four considered models.", out.width = '100%', fig.align='center'}
library("ingredients")
fifa_feat_gbm_shallow <- ingredients::feature_importance(fifa_gbm_exp_shallow)
#plot(fifa_feat_gbm_shallow, max_vars = 12)

fifa_feat_gbm_deep <- ingredients::feature_importance(fifa_gbm_exp_deep)
#plot(fifa_feat_gbm_deep, max_vars = 12)

fifa_feat_rf <- ingredients::feature_importance(fifa_rf_exp)
#plot(fifa_feat_rf, max_vars = 12)

fifa_feat_rms <- ingredients::feature_importance(fifa_rms_exp)
#plot(fifa_feat_rms, max_vars = 12)

plot(fifa_feat_rf, fifa_feat_rms, 
     fifa_feat_gbm_deep, fifa_feat_gbm_shallow, 
     max_vars = 12) +facet_wrap(~label, ncol = 2, scales = "free_y")
```

Figure \@ref(fig:usecaseFIFApdp) shows Partial Dependency profiles for the most important variables. They show average relation between particular variable and players value.

The general direction of relation in all models is the same. The larger the player characteristic the higher is the price. With a single exception – variable Age.

Random forest model has smallest range of average model responses. All tree-based models stabilize average predictions at the ends of variables ranges.


The most interesting difference between Exploratory Data Analysis presented in Figure \@ref(fig:distFIFA19scatter) and Exploratory Model Analysis presented in Figure \@ref(fig:usecaseFIFApdp) is related with variable `Age`. In Figure \@ref(fig:distFIFA19scatter) the relation was non-monotonic while in Figure \@ref(fig:usecaseFIFApdp) its monotonically decreasing. 
How we can explain this difference? 
One explanation is following: Youngest players have lower values not because of their age but because of lower skills that are correlated with Age. The EDA analysis cannot entangle these effects, thus for youngest players we see lover values also because their lower skills. But models learned that once we take skills into account, the effect of age is only decreasing. 

This example also shows, that proper *exploration of models may be more insightful than exploration of raw data*. Variable `Age` is correlated with other confounding variables. This entangle was visible in the EDA analysis. But models learned to disentangle these effects.


```{r usecaseFIFApdp, warning=FALSE, message=FALSE, echo=FALSE, fig.width=9, fig.height=8, fig.cap="(fig:usecaseFIFApdp) Partial dependency profiles for four selected variables.", out.width = '90%', fig.align='center'}
fifa19_pd_shallow <- ingredients::partial_dependency(fifa_gbm_exp_shallow, variables = c("Age", "Reactions","BallControl", "Dribbling"))

fifa19_pd_deep <- ingredients::partial_dependency(fifa_gbm_exp_deep, variables = c("Age", "Reactions","BallControl", "Dribbling"))

fifa19_pd_rf <- ingredients::partial_dependency(fifa_rf_exp, variables = c("Age", "Reactions","BallControl", "Dribbling"))

fifa19_pd_rms <- ingredients::partial_dependency(fifa_rms_exp, variables = c("Age", "Reactions","BallControl", "Dribbling"))

plot(fifa19_pd_shallow, fifa19_pd_deep, fifa19_pd_rf, fifa19_pd_rms) +
  scale_y_log10() +
  ggtitle("Partial Dependency profiles for selected variables")
```

## Instance understanding

Time to see how the model behaves for a single observation / player 
This can be done for any player, but for this example we will use *Robert Lewandowski*, the most valuable polish football player.

Here are his characteristics in the FIFA 19 database.

```{r RobertLewandowski}
fifa19small["R. Lewandowski",]
```

In the chapter \@ref(breakDown) we showed a Break Down plots for presentation of variable attributions. In the Figure \@ref(fig:usecaseFIFAbreakDown) we show Break Down plots for Robert Lewandowski predictions.

```{r usecaseFIFAbreakDown, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=7, fig.cap="(fig:usecaseFIFAbreakDown) Break down plot for GBM model.", out.width = '70%', fig.align='center'}
library("iBreakDown")
fifa_pg <- break_down(fifa_gbm_exp_shallow, new_observation = fifa19small["R. Lewandowski",])
plot(fifa_pg) +
  ggtitle("Break Down plot for Robert Lewandowski (GBM model)")
```

In the chapter \@ref(shapley) we showed a SHAP values for presentation of variable attributions. In the Figure \@ref(fig:usecaseFIFAshap) we show SHAP plots for Robert Lewandowski predictions. As it was expected, these explanations are consistent.

```{r usecaseFIFAshap, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=7, fig.cap="(fig:usecaseFIFASHAP) SHAP values for GBM model.", out.width = '70%', fig.align='center'}
fifa_pg <- shap(fifa_gbm_exp_shallow, new_observation = fifa19small["R. Lewandowski",])
plot(fifa_pg) +
  ggtitle("SHAP values plot for Robert Lewandowski (GBM model)")
```


Robert Lewandowski is a striker. It makes sense that his most valuable characteristics are Reactions and BallControl.

How these plots will look like for goalkeepers? Figure \@ref(fig:usecaseFIFAbreakDownWS) show Break Down plots for Wojciech Szczęsny - most valuable polish goal keeper. As we see the most important coefficients make sense, most of them are liked with properties of goalkeepers.


```{r usecaseFIFAbreakDownWS, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=7, fig.cap="(fig:usecaseFIFAbreakDownWS) Break down plot for GBM model.", out.width = '70%', fig.align='center'}
fifa_pg <- break_down(fifa_gbm_exp_shallow, new_observation = fifa19small["W. Szczęsny",])
plot(fifa_pg) +
  ggtitle("Break Down plot for Wojciech Szczęsny (GBM model)")
```


In chapter \@ref(ceterisParibus) we introduced Ceteris Paribus profiles. These are more details steps of the model exploration. Based on an example of Robert Lewandowski, let's see how change in one characteristic affects model value.

All models give Robert best scores when it comes to `Reactions`, `BallControl` or `Dribling`. When it comes to Age we see that the predicted value is just before a larger drop in value prediction.


```{r usecaseFIFAceterisParibus, warning=FALSE, message=FALSE, echo=FALSE, fig.width=8, fig.height=6.5, fig.cap="(fig:usecaseFIFAceterisParibus) Ceteris Paribus plots for four selected observations.", out.width = '90%', fig.align='center'}

fifa_cp_shallow <- ceteris_paribus(fifa_gbm_exp_shallow,
                           new_observation = fifa19small["R. Lewandowski",], variables = c("Age", "Reactions","BallControl", "Dribbling"),
                           variable_splits = list(Age = seq(15,45,0.1), Reactions = seq(20,100,0.1), BallControl = seq(20,100,0.1), Dribbling = seq(20,100,0.1))
                           )

fifa_cp_deep <- ceteris_paribus(fifa_gbm_exp_deep,
                           new_observation = fifa19small["R. Lewandowski",], variables = c("Age", "Reactions","BallControl", "Dribbling"),
                           variable_splits = list(Age = seq(15,45,0.1), Reactions = seq(20,100,0.1), BallControl = seq(20,100,0.1), Dribbling = seq(20,100,0.1))
                           )

fifa_cp_rf <- ceteris_paribus(fifa_rf_exp,
                           new_observation = fifa19small["R. Lewandowski",], variables = c("Age", "Reactions","BallControl", "Dribbling"),
                           variable_splits = list(Age = seq(15,45,0.1), Reactions = seq(20,100,0.1), BallControl = seq(20,100,0.1), Dribbling = seq(20,100,0.1))
                           )

fifa_cp_rms <- ceteris_paribus(fifa_rms_exp,
                           new_observation = fifa19small["R. Lewandowski",], variables = c("Age", "Reactions","BallControl", "Dribbling"),
                           variable_splits = list(Age = seq(15,45,0.1), Reactions = seq(20,100,0.1), BallControl = seq(20,100,0.1), Dribbling = seq(20,100,0.1))
                           )

plot(fifa_cp_shallow, fifa_cp_deep, fifa_cp_rf, fifa_cp_rms, color = "_label_") + 
  show_observations(fifa_cp_rf, fifa_cp_shallow, fifa_cp_deep,fifa_cp_rms, variables = c("Age", "Reactions","BallControl", "Dribbling")) + 
  scale_y_log10()

```
