# Reproducibility

All examples presented in this book are reproducible. Parts of the source codes are available in the book. Over time, some of the functionality of the described packages may change. The online version is updated. Fully reproducible code is available at https://pbiecek.github.io/ema/. 

Possible differences may be caused by other versions of the installed packages. Results in this version of the book are obtained with the following versions of the packages.

## Package versions for R

The current versions of packages in R can be checked with `sessionInfo()`.

```{r}
sessionInfo()
```

## Package versions for Python

The current versions of packages in Python can be checked with `pip`.

```{r, eval=FALSE}
pip freeze 
```

```
 absl-py==0.6.1             appnope==0.1.0            
 astor==0.7.1               atomicwrites==1.2.1       
 attrs==18.2.0              backcall==0.1.0           
 bleach==3.1.1              cycler==0.10.0            
 dalex==0.3.0               dbexplorer==1.21          
 decorator==4.4.2           defusedxml==0.6.0         
 entrypoints==0.3           future==0.17.1            
 gast==0.2.0                grpcio==1.16.1            
 h5py==2.8.0                imageio==2.9.0            
 importlib-metadata==1.5.0  innvestigate==1.0.4       
 ipykernel==5.1.4           ipython==7.13.0           
 ipython-genutils==0.2.0    jedi==0.16.0              
 Jinja2==2.11.1             joblib==0.14.1            
 json5==0.9.2               jsonschema==3.2.0         
 jupyter-client==6.0.0      jupyter-core==4.6.3       
 jupyterlab==2.0.1          jupyterlab-server==1.0.7  
 Keras==2.2.2               Keras-Applications==1.0.6 
 Keras-Preprocessing==1.0.5 kiwisolver==1.0.1         
 lightgbm==2.3.1            lime==0.2.0.1             
 Markdown==3.0.1            MarkupSafe==1.1.1         
 matplotlib==2.2.2          mistune==0.8.4            
 more-itertools==4.3.0      nbconvert==5.6.1          
 nbformat==5.0.4            networkx==2.4             
 notebook==6.0.3            numpy==1.19.0             
 pandas==1.1.1              pandocfilters==1.4.2      
 parso==0.6.2               patsy==0.5.1              
 pd==0.0.1                  pexpect==4.8.0            
 pickleshare==0.7.5         Pillow==5.3.0             
 plotly==4.9.0              pluggy==0.8.0             
 prometheus-client==0.7.1   prompt-toolkit==3.0.3     
 protobuf==3.6.1            psycopg2==2.7.4           
 ptyprocess==0.6.0          py==1.7.0                 
 Pygments==2.5.2            PyMySQL==0.8.0            
 pyodbc==4.0.23             pyparsing==2.2.0          
 pyrsistent==0.15.7         pytest==4.0.0             
 python-dateutil==2.7.3     pytz==2018.4              
 PyWavelets==1.1.1          PyYAML==3.13              
 pyzmq==19.0.0              retrying==1.3.3           
 scikit-image==0.17.2       scikit-learn==0.22.2.post1
 scipy==1.1.0               Send2Trash==1.5.0         
 simplejson==3.14.0         six==1.11.0               
 sklearn==0.0               sphinx-rtd-theme==0.4.0   
 statsmodels==0.11.1        tabulate==0.8.7           
 tensorboard==1.12.0        tensorflow==1.12.0        
 termcolor==1.1.0           terminado==0.8.3          
 testpath==0.4.4            tifffile==2020.7.17       
 torch==1.0.1.post2         torchvision==0.2.1        
 tornado==6.0.4             tqdm==4.43.0              
 traitlets==4.3.3           typing==3.6.4             
 virtualenv==16.2.0         wcwidth==0.1.8            
 webencodings==0.5.1        Werkzeug==0.14.1          
 xgboost==0.72              zipp==3.1.0               
```

